---
title: "Data visualisation"
author: "Ihsan Khaliq"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data visualisation}
  %\VignetteEngine{knitr::rmarkdown_notangle}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
knitr::opts_chunk$set(progress = TRUE, verbose = TRUE)
knitr::opts_chunk$set(
  echo = TRUE,
  fig.width = 12,
  fig.height = 12,
  fig.align = "center"
)

devtools::load_all()
```

Load libraries

```{r message=FALSE, echo=TRUE, warning=FALSE}
library("grDevices")
library("tidyverse")
library("viridis")
library("showtext")
library("here")
library("ggpubr")
library("extrafont")
library("tidyr")
library("lubridate")
library("readxl")
library("SDMTools")
library("kableExtra")
library("clifro")
library("ragg")
library("ggthemes")
library("patchwork")
library("nasapower")
library("Metrics")
library("pROC")
library("metrica")
library("lemon")
library("ggeffects")
library("ggpubr")
library("grid")
library("epibxb")
```


```{r fonts, include=FALSE, message=FALSE, eval=FALSE, message=FALSE}
# Note that GhostScript needs to be installed at the system level for the PS files to be generated.
# MacOS users can use `brew install ghostscript`
# 
# Windows users can follow these directions:
# 1.	Go to the GhostScript website (https://www.ghostscript.com/download/gsdnld.html)
# 2.	Download the windows installer suitable for your machine
# 3.	Run the installer file which you downloaded and follow the prompts
# 4.	After running the installer click the windows "Start" button and type "Edit environment variables for your account" and open
# 5.	In the tab 'Advanced' click the button at the bottom 'Environment Variables...'
# 6.	Under 'System variables' find the variable 'Path', select 'Path' and click the 'Edit' button
# 7. 	Select a new line and copy the Ghostscript 'bin' folder location into the field.
# 7.1	If you installed ghostscript to the default folder location; then the folder location will likely be "C:\Program Files\gs\gs9.52\bin", the version number (9.52) may differ.
# 8.	Save and exit the environmental variables window
# This chunk is then run only if knitting on new computer that the files have not been generated on
# this is necessary to embed fonts in .eps files 
library("extrafont")
if (.Platform$OS.type == "windows") {
   font_import(pattern = "arial", prompt = FALSE)
   loadfonts(device = "postscript", quiet = TRUE)
} else {
   font_import(pattern = "Arial", prompt = FALSE)
   loadfonts(device = "postscript", quiet = TRUE)
}
```



### Import disease data data 

Import disease data recorded at the Lowgap & Lamsburg sites 

```{r import-disease-data, message=FALSE, warning=FALSE}
disease_dat <- read_excel(system.file(
  "extdata",
  "NC_disease_data.xlsx",
  package = "epibxb",
  mustWork = TRUE
)) %>%
  dplyr::mutate(year = as.factor(year)) %>%
  dplyr::mutate(location = as.factor(location)) %>%
  dplyr::mutate(spread_event = as.factor(spread_event)) %>%
  dplyr::mutate(replicate = as.factor(replicate)) %>%
  dplyr::mutate(treatment = as.factor(treatment)) %>%
  dplyr::mutate(total_count = as.integer(total_count)) %>%
  dplyr::mutate(month = as.factor(months(date_in))) %>%
  dplyr::mutate(spev_duration = as.integer(difftime(date_out, date_in))) %>%
  relocate(month, .after = date_out) %>%
  na.omit() %>%
  group_by(
    year,
    location,
    spread_event,
    month,
    treatment,
    date_in,
    date_out,
    cultivar,
    spev_duration
  ) %>%
  summarise(total_count = sum(total_count))
```



### Import weather data 

```{r import-weather-data, message=FALSE, warning=FALSE}
# Filter rainy periods to calculate average wind speed, wind direction & temperature wet period
weather_dat_rain <- read_excel(system.file(
  "extdata",
  "NC_weather_data.xlsx",
  package = "epibxb",
  mustWork = TRUE
)) %>%
  select(year,
         wind_speed,
         wind_direction,
         temperature,
         precipitation,
         location,
         spread_event) %>%
  dplyr::mutate(year = as.factor(year)) %>%
  dplyr::mutate(location = as.factor(location)) %>%
  dplyr::mutate(spread_event = as.factor(spread_event)) %>%
  dplyr::mutate(rain_duration = as.integer(precipitation > 0)) %>%
  filter(precipitation > 0) %>%
  group_by(year, location, spread_event) %>%
  summarise(
    total_rain = round(sum(precipitation), 5),
    mean_ws = round(mean(wind_speed), 2),
    rain_duration = round(sum(rain_duration * 15 / 60), 2),
    mean_wd = round(circular.averaging(wind_direction), 2),
    mean_temp = round(mean(temperature), 2)
  )

# Filter rainless periods to calculate mean RH
weather_dat_no_rain <-
  read_excel(system.file(
    "extdata",
    "NC_weather_data.xlsx",
    package = "epibxb",
    mustWork = TRUE
  )) %>%
  select(
    year,
    relative_humidity,
    leaf_wetness_duration,
    precipitation,
    location,
    spread_event,
    date
  ) %>%
  dplyr::mutate(year = as.factor(year)) %>%
  dplyr::mutate(location = as.factor(location)) %>%
  dplyr::mutate(spread_event = as.factor(spread_event)) %>%
  filter(precipitation == 0) %>%
  group_by(year, location, spread_event) %>%
  summarise(mean_rh = round(mean(relative_humidity * 100), 2))

# Combine data
weather_dat_comb <-
  left_join(weather_dat_rain,
            weather_dat_no_rain,
            by = c("year", "location", "spread_event"))

# Leaf wetness duration both inside and outside rainy periods
weather_wet <- read_excel(system.file(
  "extdata",
  "NC_weather_data.xlsx",
  package = "epibxb",
  mustWork = TRUE
)) %>%
  dplyr::mutate(year = as.factor(year)) %>%
  dplyr::mutate(location = as.factor(location)) %>%
  dplyr::mutate(spread_event = as.factor(spread_event)) %>%
  group_by(year, location, spread_event) %>%
  summarise(lwd_duration = round(sum(leaf_wetness_duration / 60), 2))

weather_dat <-
  left_join(weather_dat_comb,
            weather_wet,
            by = c("year", "location", "spread_event"))

# Divide week 1 of 2014 rain/rain duration/wetness duration by 4 & that of week 2 & 3 by 3 to convert to per week data because the duration of spread event was 4 and 3 weeks, respectively.

weather_dat <- weather_dat %>%
  mutate(
    total_rain = ifelse(
      year == "2017" & spread_event == "1",
      total_rain / 4,
      ifelse(
        year == "2017" &
          spread_event %in% c("2", "3"),
        total_rain / 3,
        total_rain
      )
    ),
    rain_duration = ifelse(
      year == "2017" & spread_event == "1",
      rain_duration / 4,
      ifelse(
        year == "2017" &
          spread_event %in% c("2", "3"),
        rain_duration / 3,
        rain_duration
      )
    ),
    lwd_duration = ifelse(
      year == "2017" & spread_event == "1",
      lwd_duration / 4,
      ifelse(
        year == "2017" &
          spread_event %in% c("2", "3"),
        lwd_duration / 3,
        lwd_duration
      )
    )
  )
```



### Cobmine weather & disease data

Combine weather and disease data

```{r combine-disease-weather-data, message=FALSE, warning=FALSE}
dat_NC <-
  left_join(disease_dat,
            weather_dat,
            by = c("year", "location", "spread_event")) %>%
  # Replace NA with zero because NA are introduced due to data munging. Original values were zero
  dplyr::mutate(total_rain = replace_na(total_rain, 0)) %>%
  dplyr::mutate(rain_duration = replace_na(rain_duration, 0))

# Since we filtered data separately for precipitation and then without precipitation, NAs are introduced. In this step, data (in which values were added manually) is imported

dat_missing <- read_excel(system.file(
  "extdata",
  "NC_missing_data.xlsx",
  package = "epibxb",
  mustWork = TRUE
)) %>%
  dplyr::mutate(year = as.factor(year)) %>%
  dplyr::mutate(location = as.factor(location)) %>%
  dplyr::mutate(spread_event = as.factor(spread_event))

# Combine data to replace NA values with distinct data
dat_nc <-
  left_join(dat_NC, dat_missing, by = c("year", "location", "spread_event")) %>%
  mutate(mean_ws = coalesce(mean_ws.x, mean_ws.y)) %>%
  select(-mean_ws.x, -mean_ws.y) %>%
  mutate(mean_temp = coalesce(mean_temp.x, mean_temp.y)) %>%
  select(-mean_temp.x, -mean_temp.y) %>%
  mutate(mean_rh = coalesce(mean_rh.x, mean_rh.y)) %>%
  select(-mean_rh.x, -mean_rh.y) %>%
  mutate(mean_wd = coalesce(mean_wd.x, mean_wd.y)) %>%
  select(-mean_wd.x, -mean_wd.y) %>%
  mutate(lwd_duration = coalesce(lwd_duration.x, lwd_duration.y)) %>%
  select(-lwd_duration.x, -lwd_duration.y) %>%
  distinct()

dat_nc <- dat_nc %>%
  mutate(daily_rain = round(total_rain/spev_duration, 2),
         daily_lwd  = round(lwd_duration/spev_duration, 2))



# Filter out mulch treatment. Use non-mulch and CP only.
dat_nc_ncb <- dat_nc %>%
  filter(treatment != "mulch", treatment != "between_row") # filter non-mulch, CP and between row treatments data

# Data considering only CP treatment
dat_cp <- dat_nc %>%
  filter(treatment == "CP")

# Data considering only leaf debris treatment
dat_ld <- dat_nc %>%
  filter(treatment == "non_mulch")

# Data considering only between row treatment
dat_br <- dat_nc %>%
  filter(treatment == "between_row")

# Data for Lambsburg site only
dat_lambsburg <- dat_nc %>%
  filter(location == "Lambsburg") %>%
  filter(treatment != "mulch", treatment != "between_row")
```



### Histogram 

Histogram plot to visualize the distribution of the number of infected leaves per plant.

```{r data-distribution-histogram}
ggplot(dat_nc_ncb, aes(x = total_count)) +
  geom_histogram(binwidth = 2,
                 fill = "steelblue",
                 alpha = 0.7) +
  xlab("Number of infected leaves")
```



### Infected leaf count per week

```{r column-plot}
dat_week <- dat_nc_ncb %>%
  mutate(date_in = format(date_in, "%m/%d")) %>%
  group_by(year, date_in) %>%
  arrange(year, date_in, total_count) %>%
  ungroup() %>%
  mutate(date_in = as.factor(date_in)) %>%
  mutate(treatment = recode(treatment,
                            "CP" = "Infected canopies",
                            "non_mulch" = "Leaf debris")) %>%
  rename("Inoculum source" = treatment)

dates_weeks <- dat_week |>
  mutate(date = mdy(paste0(date_in, year)),
         week = week(date))


min_week <- min(dates_weeks$week)
max_week <- max(dates_weeks$week)

year_breaks <- 2017:2014 |>
  set_names() |>
  map(\(year_n) {
    year_dat <- dates_weeks |>
      filter(year == year_n) |>
      select(date_in, week) |>
      unique()
    
    week_map <- character(max_week - min_week + 1)
    names(week_map) <- seq(min_week, max_week)
    
    week_map[as.character(year_dat$week)] <-
      as.character(year_dat$date_in)
    
    
    week_map
    
  })


p <- dates_weeks |>
  mutate(week = factor(week),
         `Inoculum source` = factor(`Inoculum source`)) |>
  nest(data = -year) |>
  mutate(pl = map2(data, year, \(data, year) {
    # browser()
    
    ggplot(data, aes(x = week, y = total_count, fill = `Inoculum source`)) +
      geom_col() +
      scale_x_discrete(labels = year_breaks[[as.character(year)]],
                       drop = FALSE) +
      facet_wrap(year, scales = "free_y", strip.position = "right") +
      scale_fill_manual(values = c(
        "Infected canopies" = "navyblue",
        "Leaf debris" = "darkred"
      )) +
      theme_few(base_size = 10) +
      coord_cartesian(ylim = c(0, 1000))
    
  })) |>
  pull(pl) |>
  wrap_plots(ncol = 1) +
  plot_layout(guides = "collect") & xlab(NULL) & ylab(NULL) &
  theme(
    legend.position = "top",
    axis.text.x = element_text(
      angle = 90,
      hjust = 1,
      vjust = 0.5,
      color = "black"
    ),
    axis.text.y = element_text(vjust = 0.5, color = "black"),
    axis.title = element_text(color = "black"),
    strip.text = element_text(face = "bold", color = "black")
    
  )

p <-
  p + labs(x = "Date of detector plants placement in the field") +
  theme(axis.title.x = element_text(hjust = 0.5))  # Center x-axis label


p <- wrap_elements(p) +
  labs(tag = "Infected leaf count on detector plants in each monitoring week") +
  theme(plot.tag = element_text(size = rel(1), angle = 90),
        plot.tag.position = "left")
p
```



```{r}
ggsave(
  here("man", "figures/p.eps"),
  plot = p,
  width = 7,
  height = 7,
  units = "in",
  dpi = 600,
  device =  cairo_ps
)

ggsave(
  here("man", "figures/p.png"),
  plot = p,
  width = 12,
  height = 10,
  units = "in",
  dpi = 600,
  device =  cairo_ps
)
```



### Windroses showing wind speed & direction in each year

```{r windrose}
fig_S <-
  with(
    dat_nc_ncb,
    windrose(
      mean_ws,
      mean_wd,
      year,
      n_col = 2,
      legend_title = "Wind speed (m/s)"
    )
  )

fig_S3 <-
  fig_S +
  scale_fill_viridis_d(name = "Wind Speed (m/s)", direction = -1) +
  xlab("") +
  theme_pubclean(base_family = "Arial", base_size = 12)

fig_S
```



## Heatmaps

### Heatmap for leaf wetness duration & temperature relationship

```{r heatmap-temp-lwd}
dat_lwd <-
  dat_nc_ncb %>%
  as_tibble() %>%
  mutate(
    mean_temp = cut_interval(mean_temp, n = 10, dig.lab = 5),
    lwd_duration = cut_interval(lwd_duration, n = 10, dig.lab = 5),
  ) %>%
  group_by(mean_temp, lwd_duration) %>%
  summarize(total_count = sum(total_count))

# function to standardize the labels generated by cut_interval
standardize_cut_levels <- function(l, dig = 2) {
  sapply(l, \(s) {
    m = gregexpr("\\d+\\.?\\d{0,}", s)
    m = regmatches(s, m)[[1]]
    r = sub(m[1], trimws(sprintf(
      paste0("%9.", dig, "f"), as.numeric(m[1])
    )), s)
    sub(m[2], trimws(sprintf(
      paste0("%9.", dig, "f"), as.numeric(m[2])
    )), r)
  })
}


levels(dat_lwd$mean_temp) <-
  standardize_cut_levels(levels(dat_lwd$mean_temp))
levels(dat_lwd$lwd_duration) <-
  standardize_cut_levels(levels(dat_lwd$lwd_duration))


heatmap_lwd <- ggplot(dat_lwd, aes(mean_temp, lwd_duration)) +
  geom_tile(aes(fill = total_count), color = "black") +
  geom_text(aes(label = round(total_count, 1))) +
  scale_fill_gradient(low = "white", high = "red") +
  labs(x = "Mean temperature (°C)", y = "Leaf wetness duration (h)") +
  guides(fill = "none") +
  theme(panel.grid.major = element_blank()) +
  theme_few(base_size = 12, base_family = "Arial")

heatmap_lwd
```


```{r save, include=FALSE, eval=TRUE}
ggsave(here::here("man", "figures/heatmap_lwd.png"))
ggsave(here::here("man", "figures/heatmap_lwd.eps"), device = cairo_ps)
```


### Heatmap for rain & temperature relationship

```{r heatmap-temp-rain}
dat_rain <-
  dat_nc_ncb %>%
  as_tibble() %>%
  mutate(
    mean_temp = cut_interval(mean_temp, n = 10, dig.lab = 5),
    total_rain = cut_interval(total_rain, n = 10, dig.lab = 5),
  ) %>%
  group_by(mean_temp, total_rain) %>%
  summarize(total_count = sum(total_count))

# function to standardize the labels generated by cut_interval
standardize_cut_levels <- function(l, dig = 2) {
  sapply(l, \(s) {
    m = gregexpr("\\d+\\.?\\d{0,}", s)
    m = regmatches(s, m)[[1]]
    r = sub(m[1], trimws(sprintf(
      paste0("%9.", dig, "f"), as.numeric(m[1])
    )), s)
    sub(m[2], trimws(sprintf(
      paste0("%9.", dig, "f"), as.numeric(m[2])
    )), r)
  })
}

levels(dat_rain$mean_temp) <-
  standardize_cut_levels(levels(dat_rain$mean_temp))
levels(dat_rain$total_rain) <-
  standardize_cut_levels(levels(dat_rain$total_rain))

heatmap_rain <- ggplot(dat_rain, aes(mean_temp, total_rain)) +
  geom_tile(aes(fill = total_count), color = "black") +
  geom_text(aes(label = round(total_count, 1))) +
  scale_fill_gradient(low = "white", high = "red") +
  labs(x = "Mean temperature during rainy periods (°C)", y = "Total rain (mm)") +
  guides(fill = "none") +
  theme(panel.grid.major = element_blank()) +
  theme_few(base_size = 12, base_family = "Arial")

heatmap_rain
```


```{r save-heatmap-temp-rain, include=FALSE, eval=TRUE}
ggsave(here::here("man", "figures/heatmap_rain.png"))
ggsave(here::here("man", "figures/heatmap_rain.eps"), device = cairo_ps)
```



### Heatmap for RH & temperature relationship

```{r heatmap-temp-rh}
dat_RH <-
  dat_nc_ncb%>%
  as_tibble() %>%
  mutate(
    mean_temp = cut_interval(mean_temp, n = 10),
    mean_rh = cut_interval(mean_rh, n = 10),
  ) %>%
  group_by(mean_temp, mean_rh) %>%
  summarize(total_count = sum(total_count))


# function to standardize the labels generated by cut_interval
standardize_cut_levels <- function(l, dig=2) {
  sapply(l, \(s) {
    m = gregexpr("\\d+\\.?\\d{0,}",s)
    m = regmatches(s,m)[[1]]
    r = sub(m[1], trimws(sprintf(paste0("%9.", dig, "f"), as.numeric(m[1]))),s)
    sub(m[2], trimws(sprintf(paste0("%9.", dig, "f"), as.numeric(m[2]))),r)
  })
}


levels(dat_RH$mean_temp) <- standardize_cut_levels(levels(dat_RH$mean_temp))
levels(dat_RH$mean_rh) <- standardize_cut_levels(levels(dat_RH$mean_rh))


heatmap_rh <- ggplot(dat_RH, aes(mean_temp, mean_rh)) +
  geom_tile(aes(fill = total_count), color = "black") +
  geom_text(aes(label = round(total_count, 1))) +
  scale_fill_gradient(low = "white", high = "red") +
  labs(x = "Mean temperature (°C)", y = "Mean relative humidity (%)") +
  guides(fill = "none") + 
  theme(panel.grid.major = element_blank()) +
  theme_few(base_size = 12, base_family = "Arial") 

heatmap_rh
```


```{r save-heatmap-rh-temp, include=FALSE, eval=TRUE}
ggsave(here::here("man", "figures/heatmap_rh.png"))
ggsave(here::here("man", "figures/heatmap_rh.eps"), device = cairo_ps)
```



### Combined rain calendar heatmap (Lambsburg + Lowgap)

```{r calendar-heatmap-rain, warning=FALSE}
df1 <-
  read_excel(system.file(
    "extdata",
    "NC_weather_data.xlsx",
    package = "epibxb",
    mustWork = TRUE
  )) %>%
  dplyr::mutate(date = as.Date(date, "%m/%d/%Y")) %>%
  dplyr::mutate(year = as.factor(year)) %>%
  group_by(date, year) %>%
  summarise(precipitation = round(sum(precipitation))) %>%
  ungroup()


# color ramp
pubu <- RColorBrewer::brewer.pal(9, "PuBu")
col_p <- colorRampPalette(pubu)

theme_calendar <- function() {
  theme(
    aspect.ratio = 1 / 2,
    
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text.y = element_blank(),
    axis.text = element_text(),
    
    panel.grid = element_blank(),
    panel.background = element_blank(),
    
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 15),
    
    legend.position = "top",
    legend.text = element_text(hjust = .5),
    legend.title = element_text(size = 9, hjust = 1),
    
    plot.caption =  element_text(hjust = 1, size = 8),
    panel.border = element_rect(
      colour = "grey",
      fill = NA,
      size = 1
    ),
    plot.title = element_text(
      hjust = .5,
      size = 26,
      face = "bold",
      margin = margin(0, 0, 0.5, 0, unit = "cm")
    ),
    plot.subtitle = element_text(hjust = .5, size = 16)
  )
}



dat_prr <- df1 %>%
  rename(pr = precipitation) %>%
  complete(date = seq(min(date),
                      max(date),
                      "day")) %>%
  mutate(
    weekday = lubridate::wday(date, label = T, week_start = 1),
    month = lubridate::month(date, label = T, abbr = F),
    week = isoweek(date),
    day = day(date)
  ) %>%
  na.omit()

dat_prr <- mutate(
  dat_prr,
  week = case_when(
    month == "December" & week == 1 ~ 53,
    month == "January" &
      week %in% 52:53 ~ 0,
    TRUE ~ week
  ),
  pcat = cut(pr, c(-1, 0, 0.5, 1:5, 7, 9, 25, 81)),
  text_col = ifelse(pcat %in% c("(7,9]", "(9,25]"), "white", "black")
)

calendar_combined_rain <- ggplot(dat_prr %>%
                                   mutate(week = week - min(week),
                                          .by = c(year, month)),
                                 aes(weekday, -week, fill = pcat)) +
  geom_tile(colour = "white", size = .4)  +
  geom_text(aes(label = day, colour = text_col), size = 2.5) +
  guides(fill = guide_colorsteps(
    barwidth = 25,
    barheight = .4,
    title.position = "top"
  )) +
  scale_fill_manual(
    values = c("white", col_p(13)),
    na.value = "grey90",
    drop = FALSE
  ) +
  scale_colour_manual(values = c("black", "white"), guide = "none") +
  facet_grid(year ~ month,
             scales = "free") +
  labs(title = " Daily rainfall during field trials (2014-2017)",
       subtitle = "Rainfall",
       fill = "mm") +
  theme_calendar()

calendar_combined_rain
```


```{r save-calendar-heatmap-rain}
ggsave(here::here("man", "figures/calendar_combined_rain.png"))
ggsave(here::here("man", "figures/calendar_combined_rain.eps"), device = cairo_ps)

```


### Combined temperature heatmap (Lambsburg + Lowgap)

```{r calendar-heatmap-rain-temp-Lambsburg-Lowgap, warning=FALSE}
df2 <-
  read_excel(system.file(
    "extdata",
    "NC_weather_data.xlsx",
    package = "epibxb",
    mustWork = TRUE
  )) %>%
  dplyr::mutate(date = as.Date(date, "%m/%d/%Y")) %>%
  dplyr::mutate(year = as.factor(year)) %>%
  filter(precipitation > 0) %>%
  group_by(date, year) %>%
  summarise(temperature = mean(temperature)) %>%
  ungroup()

dat_tr <- df2 %>%
  rename(tr = temperature) %>%
  complete(date = seq(min(date),
                      max(date),
                      "day")) %>%
  mutate(
    weekday = lubridate::wday(date, label = T, week_start = 1),
    month = lubridate::month(date, label = T, abbr = F),
    week = isoweek(date),
    day = day(date)
  ) %>%
  na.omit()

dat_tr <- mutate(
  dat_tr,
  week = case_when(
    month == "December" & week == 1 ~ 53,
    month == "January" &
      week %in% 52:53 ~ 0,
    TRUE ~ week
  ),
  pcat = cut(tr, c(1.8, 3:7, 9, 15, 22, 29)),
  text_col = ifelse(pcat %in% c("(9,15]", "(15,22]"), "white", "black")
)


calendar_combined_temp <- ggplot(dat_tr %>%
                                   mutate(week = week - min(week),
                                          .by = c(year, month)),
                                 aes(weekday,-week, fill = tr)) +
  geom_tile(colour = "white", linewidth = 0.4) +
  geom_text(aes(label = day), size = 2.5, colour = "black") + # Set text color to black
  guides(fill = guide_colorsteps(
    barwidth = 25,
    barheight = 0.4,
    title.position = "top"
  )) +
  scale_fill_steps2(
    low = "blue",
    high = "red",
    midpoint = 22,
    n.breaks = 10,
    # Set the breaks to match the custom labels
    breaks = c(3, 5, 8, 12, 15, 18, 20, 22, 25, 27),
    # Use the labels argument to specify custom labels without decimal places
    labels = c(3, 5, 8, 12, 15, 18, 20, 22, 25, 27)
  ) +
  facet_grid(year ~ month,
             scales = "free") +
  labs(title = "Daily mean temperature during rainy periods in field trials (2014-2017)",
       subtitle = "Temperature",
       fill = "degree Celsius") +
  theme_calendar()

calendar_combined_temp
```


```{r save-calendar-heatmap-rain-temp-Lambsburg-Lowgap}
ggsave(here::here("man", "figures/calendar_combined_temp.png"))
ggsave(here::here("man", "figures/calendar_combined_temp.eps"), device = cairo_ps)
```



## Number of hours per week with relative humidity greater than 90% and mean temperature when relative humidity was greater than 90%

###  During rainy periods 

```{r rh90-temp-rainy-periods}
weather_rh <-
  read_excel(system.file(
    "extdata",
    "NC_weather_data.xlsx",
    package = "epibxb",
    mustWork = TRUE
  )) %>%
  mutate(day = day(date)) %>%
  mutate(month = month(date)) %>%
  mutate(relative_humidity = relative_humidity * 100) %>%
  mutate(
    spread_event =  as.factor(spread_event),
    year = as.factor(year),
    location = as.factor(location)
  ) %>%
  mutate(lw_rh = case_when(relative_humidity > 90 ~ 1,
                           TRUE ~ 0)) %>%
  filter(lw_rh > 0) %>%
  group_by(year, location, spread_event) %>%
  summarise(temp_rh_lwd = mean(temperature),
            # mean temperature when RH was > 90%
            lw_rh_count = n() * 15 / 60) # Weekly number of hours when RH was greater than 90%


# Divide week 1 of 2017 rain/rain duration/wetness duration by 4 & that of week 2 & 3 by 3 because the duration of spread event was 4 and 3 weeks, respectively.

weather_rh <- weather_rh %>%
  mutate(
    lw_rh_count = ifelse(
      year == "2017" & spread_event == "1",
       lw_rh_count/4,
      ifelse(
        year == "2017" &
          spread_event %in% c("2", "3"),
         lw_rh_count/3,
         lw_rh_count 
      )
    ))


disease_dat2 <- disease_dat %>%
 filter(treatment != "mulch", treatment != "between_row") %>% 
  group_by(year, location, spread_event, spev_duration) %>%
  summarise(total_count = sum(total_count))
  

# Combine weather data with disease data
disease_dat_rh <- left_join(disease_dat2, weather_rh, by = c("year", "location", "spread_event")) %>%
  mutate(daily_lw_rh_count = lw_rh_count/spev_duration)


  
# openxlsx::write.xlsx(disease_dat_rh, "rh_dat_rainy_period.xlsx")
```



### Heatmap for hours with RH >90% & mean temperature during RH >90% 

```{r heatmap-temp-rh90, rainy-periods}
dat_RH2 <-
  disease_dat_rh %>%
  as_tibble() %>%
  mutate(
    temp_rh_lwd= cut_interval(temp_rh_lwd, n = 10),
    lw_rh_count = cut_interval(lw_rh_count, n = 10),
  ) %>%
  group_by(temp_rh_lwd, lw_rh_count) %>%
  summarize(total_count = sum(total_count))


# function to standardize the labels generated by cut_interval
standardize_cut_levels <- function(l, dig=2) {
  sapply(l, \(s) {
    m = gregexpr("\\d+\\.?\\d{0,}",s)
    m = regmatches(s,m)[[1]]
    r = sub(m[1], trimws(sprintf(paste0("%9.", dig, "f"), as.numeric(m[1]))),s)
    sub(m[2], trimws(sprintf(paste0("%9.", dig, "f"), as.numeric(m[2]))),r)
  })
}


levels(dat_RH2$temp_rh_lwd) <- standardize_cut_levels(levels(dat_RH2$temp_rh_lwd))
levels(dat_RH2$lw_rh_count) <- standardize_cut_levels(levels(dat_RH2$lw_rh_count))


heatmap_rh_lw <- ggplot(dat_RH2, aes(temp_rh_lwd, lw_rh_count)) +
  geom_tile(aes(fill = total_count), color = "black") +
  geom_text(aes(label = round(total_count, 1))) +
  scale_fill_gradient(low = "white", high = "red") +
  labs(x = "Mean temperature (°C)", y = "Hours relative humidity exceeding 90%") +
  guides(fill = "none") + 
  theme(panel.grid.major = element_blank()) +
  theme_few(base_size = 12, base_family = "Arial") 

heatmap_rh_lw
```


```{r save-heatmap-rh-90, include=FALSE, eval=TRUE}
ggsave(here::here("man", "figures/heatmap_rh_lw.png"))
ggsave(here::here("man", "figures/heatmap_rh_lw.eps"), device = cairo_ps)
```



## Visualise weather data for the past 12-years to help interpret temporal patterns of boxwood blight development in Virginia

### import weather data

```{r weather-data-VA}
weather_va <- get_power(
      community = "ag",
      lonlat = c(-80.7666, 36.5833),
      pars = c("RH2M", "T2M", "PRECTOTCORR"),
      dates = c("2010-01-01", "2021-12-31"),
      temporal_api = "hourly"
    ) %>%
mutate(date = as.Date(sprintf('%04d-%02d-%02d', YEAR, MO, DY))) %>%
  mutate(month = format(date, "%B")) %>%
  mutate(julian_day = yday(date)) %>%
  filter(date > "2009-12-31") %>%
rename_with(~ "rainfall", .cols = matches("PRECTOTCORR")) %>%
  rename_with(~ "temp", .cols = matches("T2M")) %>%
  rename_with(~ "rh", .cols = matches("RH2M"))
 

#openxlsx::write.xlsx(weather_va, "dailyWEATHER.xlsx")

# openxlsx::write.xlsx(weather_va, "daily_weather2010-2020_Lambsburg.xlsx")
```


### Plot monthly mean weather data for the trial site in Virginia

```{r plot-VA}
# Get monthly data for 10 years
weather_va_monthly <- weather_va %>%
  group_by(YEAR, month) %>%
  summarise(
    total_rain = round(sum(rainfall), 2),
    mean_temp = round(mean(temp), 2),
    mean_rh = round(mean(rh), 2)
  ) 

# Convert full month names to abbreviated names
weather_va_monthly$month <- factor(weather_va_monthly$month, levels = month.name, labels = month.abb)


fig_month_va <- ggplot(weather_va_monthly, aes(x = month)) +
    geom_bar(aes(y = total_rain/4, fill = "Rainfall (mm)"), stat = "identity") +
    geom_line(aes(y = mean_temp, color = "Temperature (°C)", group = 1), 
              linetype = "dashed") +
    geom_line(aes(y = mean_rh, color = "Relative Humidity (%)", group = 1), 
              linetype = "dashed") +
    geom_point(aes(y = mean_temp, color = "Temperature (°C)"), shape = 16, size = 1.5) + # Add solid circles for temperature line
    geom_point(aes(y = mean_rh, color = "Relative Humidity (%)"), shape = 16, size = 1.5) + # Add solid circles for relative humidity line
    scale_y_continuous(
        name = "Temperature and relative humidity",
        sec.axis = sec_axis(~.*4, name = "Rainfall")
    ) +
    scale_color_manual(
        name = "",
        values = c("Temperature (°C)" = "navyblue", "Relative Humidity (%)" = "darkred")
    ) +
    scale_fill_manual(name = "", values = "darkgoldenrod") +
    labs(x = "", title = "Lambsburg, Virginia") +
    theme_few() + 
    theme(
        text = element_text(size = 14),
        axis.text.x = element_text(angle = 90, vjust = 0.5),
        legend.position = "bottom",
        plot.title = element_text(hjust = 0.5)
    ) + 
    facet_wrap(~YEAR)

fig_month_va 


ggsave(
  here("man", "figures/fig_month_va.png"),
  plot = fig_month_va,
  width = 12,
  height = 7.5,
  units = "in",
  dpi = 600
)
```


### Plot daily hours with RH >90% and mean temperature during those hours for Virginia trial site

```{r plot-rh90-temp-VA}
weather_va_daily <- weather_va %>%
      mutate(days_month = days_in_month(as.Date(paste(YEAR, MO, 1, sep = "-")))) %>%
  group_by(YEAR, month) %>% 
      mutate(LW = case_when(rh > 90 ~ 1,
                            TRUE ~ 0)) %>%
      filter(LW > 0) %>%
      summarise(Air_LWD = mean(temp),
                LWD = n()
                )
# Convert full month names to abbreviated names
weather_va_daily$month <- factor(weather_va_daily$month, levels = month.name, labels = month.abb)

fig_month_va_rh <- ggplot(weather_va_daily, aes(x = month)) +
  geom_bar(aes(y = LWD / 4, fill = "Hours relative humidity above 90%"), stat = "identity") +
  geom_line(aes(y = Air_LWD, color = "Temperature (°C)", group = 1),
            linetype = "dashed") +
  geom_point(aes(y = Air_LWD)) +
  # geom_line(aes(y = mean_rh, color = "Relative Humidity (%)", group = 1), 
  #           linetype = "dashed") +
  # geom_point(aes(y = mean_rh), shape = 18) +
  #scale_x_discrete(limits = month.name) +
  scale_y_continuous(
    name = "",
    sec.axis = sec_axis(~.*4, name = "Hours relative humidity above 90%")
  ) +
  scale_color_manual(
    name = "",
    values = c("Temperature (°C)" = "black")
  ) +
  scale_fill_manual(name = "", values = "darkgoldenrod") +
  labs(x = "", title = "Lambsburg, Virginia") +
  theme_few() + 
  theme(
    text = element_text(size = 10),
    axis.text.x = element_text(angle = 90, vjust = 0.5),
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5)
  ) + 
  facet_wrap(~YEAR)

fig_month_va_rh

ggsave(
  here("man", "figures/fig_month_va_rh.png"),
  plot = fig_month_va_rh,
  width = 9,
  height = 6,
  units = "in",
  dpi = 1000

)
```


## Visualise weather data for the past decade for the trial site in North Carolina

### Import weather data

```{r import-data-NC}
weather_nc <- get_power(
      community = "ag",
      lonlat = c(-80.8333, 36.5166),
      pars = c("RH2M", "T2M", "PRECTOTCORR"),
      dates = c("2010-01-01", "2021-12-31"),
      temporal_api = "hourly"
    ) %>%
mutate(date = as.Date(sprintf('%04d-%02d-%02d', YEAR, MO, DY))) %>%
  mutate(month = format(date, "%B")) %>%
  mutate(julian_day = yday(date)) %>%
  filter(date > "2009-12-31") %>%
rename_with(~ "rainfall", .cols = matches("PRECTOTCORR")) %>%
  rename_with(~ "temp", .cols = matches("T2M")) %>%
  rename_with(~ "rh", .cols = matches("RH2M"))
 
```


## Plot monthly mean data for the trial site in North Carolina

```{r plot-data}
# Get monthly data for 12 years
weather_nc_monthly <- weather_nc %>%
  group_by(YEAR, month) %>%
  summarise(
    total_rain = round(sum(rainfall), 2),
    mean_temp = round(mean(temp), 2),
    mean_rh = round(mean(rh), 2)
  ) 

# Convert full month names to abbreviated names
weather_nc_monthly$month <- factor(weather_nc_monthly$month, levels = month.name, labels = month.abb)

fig_month_nc <- ggplot(weather_nc_monthly, aes(x = month)) +
    geom_bar(aes(y = total_rain/4, fill = "Rainfall (mm)"), stat = "identity") +
    geom_line(aes(y = mean_temp, color = "Temperature (°C)", group = 1), 
              linetype = "dashed") +
    geom_line(aes(y = mean_rh, color = "Relative Humidity (%)", group = 1), 
              linetype = "dashed") +
    geom_point(aes(y = mean_temp, color = "Temperature (°C)"), shape = 16, size = 1.5) + # Add solid circles for temperature line
    geom_point(aes(y = mean_rh, color = "Relative Humidity (%)"), shape = 16, size = 1.5) + # Add solid circles for relative humidity line
    scale_y_continuous(
        name = "Temperature and relative humidity",
        sec.axis = sec_axis(~.*4, name = "Rainfall")
    ) +
    scale_color_manual(
        name = "",
        values = c("Temperature (°C)" = "navyblue", "Relative Humidity (%)" = "darkred")
    ) +
    scale_fill_manual(name = "", values = "darkgoldenrod") +
    labs(x = "", title = "") +
    theme_few() + 
    theme(
        text = element_text(size = 14),
        axis.text.x = element_text(angle = 90, vjust = 0.5),
        legend.position = "bottom",
        plot.title = element_text(hjust = 0.5)
    ) + 
    facet_wrap(~YEAR)

fig_month_nc

ggsave(
  here("man", "figures/fig_month_nc.png"),
  plot = fig_month_nc,
  width = 12,
  height = 7,
  units = "in",
  dpi = 600
)
```



## Plot daily hours with RH >90% and mean temperature during those hours for North Carolina trial site

```{r plot-rh90-temp-NC}
weather_nc_daily <- weather_nc %>%
      mutate(days_month = days_in_month(as.Date(paste(YEAR, MO, 1, sep = "-")))) %>%
  group_by(YEAR, month) %>% 
      mutate(LW = case_when(rh > 90 ~ 1,
                            TRUE ~ 0)) %>%
      filter(LW > 0) %>%
      summarise(Air_LWD = mean(temp),
                LWD = n()
                )

# Convert full month names to abbreviated names
weather_nc_daily$month <- factor(weather_nc_daily$month, levels = month.name, labels = month.abb)

fig_month_nc_rh <- ggplot(weather_nc_daily, aes(x = month)) +
  geom_bar(aes(y = LWD/4, fill = "Hours relative humidity above 90%"), stat = "identity") +
  geom_line(aes(y =Air_LWD, color = "Temperature (°C)", group = 1), 
            linetype = "dashed") +
  geom_point(aes(y = Air_LWD)) +
  # geom_line(aes(y = mean_rh, color = "Relative Humidity (%)", group = 1), 
  #           linetype = "dashed") +
  # geom_point(aes(y = mean_rh), shape = 18) +
  #scale_x_discrete(limits = month.name) +
  scale_y_continuous(
    name = "",
    sec.axis = sec_axis(~.*4, name = "Hours relative humidity above 90%")
  ) +
  scale_color_manual(
    name = "",
    values = c("Temperature (°C)" = "black")
  ) +
  scale_fill_manual(name = "", values = "darkgoldenrod") +
  labs(x = "", title = "Lowgap, North Carolina") +
  theme_few() + 
  theme(
    text = element_text(size = 10),
    axis.text.x = element_text(angle = 90, vjust = 0.5),
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5)
  ) + 
  facet_wrap(~YEAR)

fig_month_nc_rh

ggsave(
  here("man", "figures/fig_month_nc_rh.eps"),
  plot = fig_month_nc_rh,
  width = 9,
  height = 6,
  units = "in",
  dpi = 600,
  device =  cairo_ps
)

```

