---
title: "Fit GLMMs"
author: "Ihsan Khaliq"
date: "`r Sys.Date()`"
output:
   rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fit GLMMs}
  %\VignetteEngine{knitr::rmarkdown_notangle}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
knitr::opts_chunk$set(progress = TRUE, verbose = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(
  echo = TRUE,
  fig.width = 7,
  fig.height = 7,
  fig.align = "center"
)

devtools::load_all()
```


Load libraries 

```{r libraries, echo=TRUE, warning=FALSE}
library("readxl")
library("tidyverse")
library("dplyr")
library("ggeffects")
library("here")
library("glmmTMB")
library("lubridate")
library("ggpubr")
library("kableExtra")
library("DHARMa")
library("SDMTools")
library("ggthemes")
library("patchwork")
library("epiboxwoodblight")
```



```{r fonts, include=FALSE, message=FALSE, eval=FALSE, message=FALSE}
# Note that GhostScript needs to be installed at the system level for the PS files to be generated.
# MacOS users can use `brew install ghostscript`
# 
# Windows users can follow these directions:
# 1.	Go to the GhostScript website (https://www.ghostscript.com/download/gsdnld.html)
# 2.	Download the windows installer suitable for your machine
# 3.	Run the installer file which you downloaded and follow the prompts
# 4.	After running the installer click the windows "Start" button and type "Edit environment variables for your account" and open
# 5.	In the tab 'Advanced' click the button at the bottom 'Environment Variables...'
# 6.	Under 'System variables' find the variable 'Path', select 'Path' and click the 'Edit' button
# 7. 	Select a new line and copy the Ghostscript 'bin' folder location into the field.
# 7.1	If you installed ghostscript to the default folder location; then the folder location will likely be "C:\Program Files\gs\gs9.52\bin", the version number (9.52) may differ.
# 8.	Save and exit the environmental variables window
# This chunk is then run only if knitting on new computer that the files have not been generated on
# this is necessary to embed fonts in .eps files 
library("extrafont")
if (.Platform$OS.type == "windows") {
   font_import(pattern = "arial", prompt = FALSE)
   loadfonts(device = "postscript", quiet = TRUE)
} else {
   font_import(pattern = "Arial", prompt = FALSE)
   loadfonts(device = "postscript", quiet = TRUE)
}
```



## Import disease data 

### Import disease data

```{r import-disease-data, message=FALSE, warning=FALSE}
disease_dat <- read_excel(system.file(
  "extdata",
  "NC_disease_data.xlsx",
  package = "epiboxwoodblight",
  mustWork = TRUE
)) %>%
  dplyr::mutate(year = as.factor(year)) %>%
  dplyr::mutate(location = as.factor(location)) %>%
  dplyr::mutate(spread_event = as.factor(spread_event)) %>%
  dplyr::mutate(replicate = as.factor(replicate)) %>%
  dplyr::mutate(treatment = as.factor(treatment)) %>%
  dplyr::mutate(total_count = as.integer(total_count)) %>%
  dplyr::mutate(month = as.factor(months(date_in))) %>%
  dplyr::mutate(spev_duration = as.integer(difftime(date_out, date_in))) %>%
  relocate(month, .after = date_out) %>%
  na.omit() %>%
  group_by(
    year,
    location,
    spread_event,
    month,
    treatment,
    date_in,
    date_out,
    cultivar,
    spev_duration
  ) %>%
  summarise(total_count = sum(total_count))

#openxlsx::write.xlsx(disease_dat, "Table S1.xlsx", rowNames=FALSE)
```



### Import weather data 

```{r import-weather-data, message=FALSE, warning=FALSE}
# Filter rainy periods to calculate average wind speed, wind direction & temperature wet period
weather_dat_rain <- read_excel(system.file(
  "extdata",
  "NC_weather_data.xlsx",
  package = "epiboxwoodblight",
  mustWork = TRUE
)) %>%
  select(year,
         wind_speed,
         wind_direction,
         temperature,
         precipitation,
         location,
         spread_event) %>%
  dplyr::mutate(year = as.factor(year)) %>%
  dplyr::mutate(location = as.factor(location)) %>%
  dplyr::mutate(spread_event = as.factor(spread_event)) %>%
  dplyr::mutate(rain_duration = as.integer(precipitation > 0)) %>%
  filter(precipitation > 0) %>%
  group_by(year, location, spread_event) %>%
  summarise(
    total_rain = round(sum(precipitation), 5),
    mean_ws = round(mean(wind_speed), 2),
    rain_duration = round(sum(rain_duration * 15 / 60), 2),
    mean_wd = round(circular.averaging(wind_direction), 2),
    mean_temp = round(mean(temperature), 2)
  )

# Filter rainless periods to calculate mean RH
weather_dat_no_rain <-
  read_excel(system.file(
    "extdata",
    "NC_weather_data.xlsx",
    package = "epiboxwoodblight",
    mustWork = TRUE
  )) %>%
  select(
    year,
    relative_humidity,
    leaf_wetness_duration,
    precipitation,
    location,
    spread_event,
    date
  ) %>%
  dplyr::mutate(year = as.factor(year)) %>%
  dplyr::mutate(location = as.factor(location)) %>%
  dplyr::mutate(spread_event = as.factor(spread_event)) %>%
  filter(precipitation == 0) %>%
  group_by(year, location, spread_event) %>%
  summarise(mean_rh = round(mean(relative_humidity * 100), 2))

# Combine data
weather_dat_comb <-
  left_join(weather_dat_rain,
            weather_dat_no_rain,
            by = c("year", "location", "spread_event"))

# Leaf wetness duration both inside and outside rainy periods
weather_wet <- read_excel(system.file(
  "extdata",
  "NC_weather_data.xlsx",
  package = "epiboxwoodblight",
  mustWork = TRUE
)) %>%
  dplyr::mutate(year = as.factor(year)) %>%
  dplyr::mutate(location = as.factor(location)) %>%
  dplyr::mutate(spread_event = as.factor(spread_event)) %>%
  group_by(year, location, spread_event) %>%
  summarise(lwd_duration = round(sum(leaf_wetness_duration / 60), 2))

weather_dat <-
  left_join(weather_dat_comb,
            weather_wet,
            by = c("year", "location", "spread_event"))

# Divide week 1 of 2014 rain/rain duration/wetness duration by 4 & that of week 2 & 3 by 3 to convert to per week data because the duration of spread event was 4 and 3 weeks, respectively.

weather_dat <- weather_dat %>%
  mutate(
    total_rain = ifelse(
      year == "2017" & spread_event == "1",
      total_rain / 4,
      ifelse(
        year == "2017" &
          spread_event %in% c("2", "3"),
        total_rain / 3,
        total_rain
      )
    ),
    rain_duration = ifelse(
      year == "2017" & spread_event == "1",
      rain_duration / 4,
      ifelse(
        year == "2017" &
          spread_event %in% c("2", "3"),
        rain_duration / 3,
        rain_duration
      )
    ),
    lwd_duration = ifelse(
      year == "2017" & spread_event == "1",
      lwd_duration / 4,
      ifelse(
        year == "2017" &
          spread_event %in% c("2", "3"),
        lwd_duration / 3,
        lwd_duration
      )
    )
  )
```


### Cobmine weather & disease data

Combine weather and disease data

```{r combine-data}
dat_NC <-
  left_join(disease_dat,
            weather_dat,
            by = c("year", "location", "spread_event")) %>%
  # Replace NA with zero because NA are introduced due to data munging. Original values were zero
  dplyr::mutate(total_rain = replace_na(total_rain, 0)) %>%
  dplyr::mutate(rain_duration = replace_na(rain_duration, 0))

# Since we filtered data separately for precipitation and then without precipitation, NAs are introduced. In this step, data (in which values were added manually) is imported

dat_missing <- read_excel(system.file(
  "extdata",
  "NC_missing_data.xlsx",
  package = "epiboxwoodblight",
  mustWork = TRUE
)) %>%
  dplyr::mutate(year = as.factor(year)) %>%
  dplyr::mutate(location = as.factor(location)) %>%
  dplyr::mutate(spread_event = as.factor(spread_event))

# Combine data to replace NA values with distinct data
dat_nc <-
  left_join(dat_NC, dat_missing, by = c("year", "location", "spread_event")) %>%
  mutate(mean_ws = coalesce(mean_ws.x, mean_ws.y)) %>%
  select(-mean_ws.x, -mean_ws.y) %>%
  mutate(mean_temp = coalesce(mean_temp.x, mean_temp.y)) %>%
  select(-mean_temp.x, -mean_temp.y) %>%
  mutate(mean_rh = coalesce(mean_rh.x, mean_rh.y)) %>%
  select(-mean_rh.x, -mean_rh.y) %>%
  mutate(mean_wd = coalesce(mean_wd.x, mean_wd.y)) %>%
  select(-mean_wd.x, -mean_wd.y) %>%
  mutate(lwd_duration = coalesce(lwd_duration.x, lwd_duration.y)) %>%
  select(-lwd_duration.x, -lwd_duration.y) %>%
  distinct()

dat_nc <- dat_nc %>%
  mutate(daily_rain = round(total_rain/spev_duration, 2),
         daily_lwd  = round(lwd_duration/spev_duration, 2))



# Filter out mulch treatment. Use non-mulch and CP only.
dat_nc_ncb <- dat_nc %>%
  filter(treatment != "mulch", treatment != "between_row") # filter non-mulch, CP and between row treatments data

# Data considering only CP treatment
dat_cp <- dat_nc %>%
  filter(treatment == "CP")

# Data considering only leaf debris treatment
dat_ld <- dat_nc %>%
  filter(treatment == "non_mulch")

# Data considering only between row treatment
dat_br <- dat_nc %>%
  filter(treatment == "between_row")

# Data for Lambsburg site only
dat_lambsburg <- dat_nc %>%
  filter(location == "Lambsburg") %>%
  filter(treatment != "mulch", treatment != "between_row")
```



### Check data 

```{r kable-data}
kable(dat_nc_ncb,
      format = "html",
      table.attr = "class='table table-hover'")
```


Use `set.seed()` for reproducibility purposes.

```{r set-seed, echo=TRUE}
set.seed(42)
```



## Fit univariate/bivariate glmms

### Mod_1 (Total rain)

```{r mod-1}
mod_1 <-
  glmmTMB(total_count ~ total_rain +  (1 | spread_event),
          family = nbinom2,
          data = dat_nc_ncb)

summary(mod_1)
```


### Mod_2 (Wind speed)


```{r mod-2}
mod_2 <-
  glmmTMB(total_count ~ mean_wd * mean_ws + (1 | spread_event),
          family = nbinom2,
          data = dat_nc_ncb)

summary(mod_2)
```


### Mod_3 (Wind speed)

```{r mod-3}
mod_3 <-
  glmmTMB(total_count ~ mean_wd + (1 | spread_event),
          family = nbinom2,
          data = dat_nc_ncb)

summary(mod_3)
```


### Mod_4 (Leaf wetness duration)

```{r mod-4}
mod_4 <-
  glmmTMB(total_count ~  lwd_duration +  (1 | spread_event),
          family = nbinom2,
          data = dat_nc_ncb)

summary(mod_4)
```


### Mod_5 (Relative humidity)

```{r mod-5}
mod_5 <-
  glmmTMB(total_count ~  mean_rh  + (1 | spread_event),
          family = nbinom2,
          data = dat_nc_ncb)

summary(mod_5)
```


### mod_6 (Temperature)

```{r mod-6}
mod_6 <-
  glmmTMB(total_count ~  mean_temp + (1 | spread_event),
          family = nbinom2,
          data = dat_nc_ncb)

summary(mod_6)
```



### Fit multivariate glmms


```{r mod-7, warning=FALSE}
mod_7 <-
  glmmTMB(
    total_count ~ mean_rh + total_rain + mean_wd * mean_ws + mean_temp * lwd_duration + location + (1 | spread_event), family = nbinom2,
    data = dat_nc_ncb
  )

summary(mod_7) 
```



## Model Diagnostics 

### Simulate model residuals

Check if model met data assumptions, and if the model predictions can be trusted. 

```{r simulate-residuals}
simulateResiduals(mod_7, plot = T, quantreg = T)
```



### Plot model

Plot main effects of weather variables using `ggplot2`

```{r plot-main-effects}
# Relative humidity graph
p1 <- ggpredict(mod_7, "mean_rh[56:96]",
                type = "random") %>%
  as.data.frame() %>%
  rename(mean_rh = x, total_count = predicted) %>%
  mutate(mean_rh = as.numeric(as.character(mean_rh))) %>%
  ggplot(aes(mean_rh, total_count)) +
  geom_line() +
  geom_ribbon(colour = NA,
              alpha = 0.1,
              aes(ymin = conf.low, ymax = conf.high)) +
  geom_point(data = dat_nc_ncb,
             size = 1) +
  annotate("text",
           x = 80,
           y = 500,
           label = "p=0.0001") +
  # scale_color_distiller(palette = "Spectral") +
  # scale_fill_distiller(palette = "Spectral", guide = "none") +
  #coord_cartesian(ylim = range(dat_nc_ncb$total_count), xlim = range(dat_nc_ncb$mean_rh)) +
  coord_cartesian(ylim = c(0, 690)) +
  #scale_y_continuous(trans = "log1p") +
  theme_few(base_size = 11) +
  labs(x = "Mean RH (%)", y = "Number of infected leaves")

p1


# Rainfall graph
p2 <- ggpredict(mod_7, "total_rain[0:135]",
                type = "random") %>%
  as.data.frame() %>%
  rename(total_rain = x, total_count = predicted) %>%
  #mutate(total_rain= as.numeric(as.character(total_rain))) %>%
  ggplot(aes(total_rain, total_count)) +
  geom_line() +
  geom_point(data = dat_nc_ncb,
             size = 1) +
  geom_ribbon(
    colour = NA,
    fill = "black",
    alpha = 0.1,
    aes(ymin = conf.low, ymax = conf.high)
  ) +
  annotate("text",
           x = 80,
           y = 500,
           label = "p=0.0049") +
  # scale_color_distiller(palette = "Spectral") +
  # scale_fill_distiller(palette = "Spectral", guide = "none") +
  #coord_cartesian(ylim = range(dat_nc_ncb$total_count), xlim = range(dat_nc_ncb$mean_rh)) +
  coord_cartesian(ylim = c(0, 690), xlim = c(0, 135)) +
  theme_few(base_size = 11) +
  labs(x = "Total rain (mm)", y = "")

p2


fig_3 <- p1 + p2 + plot_layout(tag_level = 'new') +
  plot_annotation(tag_levels = list(c('(a)', '(b)'))) &
    theme(plot.tag = element_text(face = 'bold', size = 11))
```


```{r save-fig_3, include=FALSE, eval=TRUE}
ggsave(here::here("man", "figures/Fig_3.png"))
ggsave(here::here("man", "figures/Fig_3.eps"), device = cairo_ps)
```



### Plot interaction effect using `ggplot2`

```{r plot-interaction-effect-plot, message=FALSE, warning=FALSE}
fig_4 <- ggpredict(mod_7, terms = c("lwd_duration[1:160]", "mean_temp[10:27]"), type = "random") %>%
    as.data.frame() %>%
    rename(lwd_duration = x, mean_temp = group, total_count = predicted) %>%
    mutate(mean_temp = as.numeric(as.character(mean_temp)))
ggplot() +
    geom_line(data = fig_4, aes(lwd_duration, total_count, color = mean_temp, group = mean_temp), alpha = 0.8) +
    geom_point(data = dat_nc_ncb, aes(x = lwd_duration, y = total_count, fill = mean_temp), shape = 21, color = "black", size = 2.5) +
    scale_color_distiller(palette = "Spectral") +
    scale_fill_distiller(palette = "Spectral", guide = "none") +
   coord_cartesian(ylim=c(0, 690),xlim = c(1,160)) +
    annotate("text", x=80, y=500, label= "p=0.0001") +
    theme_few(base_size = 11) +
    labs(x = "Leaf Wetness Duration", y = "Number of Infected Leaves", color = "Mean Temperature (°C)")
```



```{r save-fig_4, include=FALSE, eval=TRUE}
ggsave(here::here("man", "figures/Fig_4.png"))
ggsave(here::here("man", "figures/Fig_4.eps"), device = cairo_ps)
```












































