---
title: "Boxwood blight infection risk model"
author: "Ihsan Khaliq"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Boxwood blight infection risk model}
  %\VignetteEngine{knitr::rmarkdown_notangle}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
knitr::opts_chunk$set(progress = TRUE, verbose = TRUE)
knitr::opts_chunk$set(
  echo = TRUE,
  fig.width = 10,
  fig.height = 10,
  fig.align = "center"
)

devtools::load_all()
```

Load libraries

```{r message=FALSE, echo=TRUE, warning=FALSE}
library("grDevices")
library("tidyverse")
library("viridis")
library("showtext")
library("here")
library("ggpubr")
library("extrafont")
library("tidyr")
library("lubridate")
library("readxl")
library("SDMTools")
library("kableExtra")
library("ragg")
library("ggthemes")
library("patchwork")
library("nasapower")
library("Metrics")
library("pROC")
library("metrica")
library("ggeffects")
library("ggpubr")
library("grid")
library("epiboxwoodblight")
```


```{r fonts, include=FALSE, message=FALSE, eval=FALSE, message=FALSE}
# Note that GhostScript needs to be installed at the system level for the PS files to be generated.
# MacOS users can use `brew install ghostscript`
# 
# Windows users can follow these directions:
# 1.	Go to the GhostScript website (https://www.ghostscript.com/download/gsdnld.html)
# 2.	Download the windows installer suitable for your machine
# 3.	Run the installer file which you downloaded and follow the prompts
# 4.	After running the installer click the windows "Start" button and type "Edit environment variables for your account" and open
# 5.	In the tab 'Advanced' click the button at the bottom 'Environment Variables...'
# 6.	Under 'System variables' find the variable 'Path', select 'Path' and click the 'Edit' button
# 7. 	Select a new line and copy the Ghostscript 'bin' folder location into the field.
# 7.1	If you installed ghostscript to the default folder location; then the folder location will likely be "C:\Program Files\gs\gs9.52\bin", the version number (9.52) may differ.
# 8.	Save and exit the environmental variables window
# This chunk is then run only if knitting on new computer that the files have not been generated on
# this is necessary to embed fonts in .eps files 
library("extrafont")
if (.Platform$OS.type == "windows") {
   font_import(pattern = "arial", prompt = FALSE)
   loadfonts(device = "postscript", quiet = TRUE)
} else {
   font_import(pattern = "Arial", prompt = FALSE)
   loadfonts(device = "postscript", quiet = TRUE)
}
```



### Import disease data data 

Import disease data recorded at the Lowgap & Lamsburg sites 

```{r import-disease-data, warning=FALSE}
disease_dat <- read_excel(system.file(
  "extdata",
  "NC_disease_data.xlsx",
  package = "epiboxwoodblight",
  mustWork = TRUE
)) %>%
  dplyr::mutate(year = as.factor(year)) %>%
  dplyr::mutate(location = as.factor(location)) %>%
  dplyr::mutate(spread_event = as.factor(spread_event)) %>%
  dplyr::mutate(replicate = as.factor(replicate)) %>%
  dplyr::mutate(treatment = as.factor(treatment)) %>%
  dplyr::mutate(total_count = as.integer(total_count)) %>%
  dplyr::mutate(month = as.factor(months(date_in))) %>%
  dplyr::mutate(spev_duration = as.integer(difftime(date_out, date_in))) %>%
  relocate(month, .after = date_out) %>%
  na.omit() %>%
  group_by(
    year,
    location,
    spread_event,
    month,
    treatment,
    date_in,
    date_out,
    cultivar,
    spev_duration
  ) %>%
  summarise(total_count = sum(total_count))
```



### Import weather data 

```{r import-weather-data, message=FALSE, warning=FALSE}
# Filter rainy periods to calculate average wind speed, wind direction & temperature wet period
weather_dat_rain <- read_excel(system.file(
  "extdata",
  "NC_weather_data.xlsx",
  package = "epiboxwoodblight",
  mustWork = TRUE
)) %>%
  select(year,
         wind_speed,
         wind_direction,
         temperature,
         precipitation,
         location,
         spread_event) %>%
  dplyr::mutate(year = as.factor(year)) %>%
  dplyr::mutate(location = as.factor(location)) %>%
  dplyr::mutate(spread_event = as.factor(spread_event)) %>%
  dplyr::mutate(rain_duration = as.integer(precipitation > 0)) %>%
  filter(precipitation > 0) %>%
  group_by(year, location, spread_event) %>%
  summarise(
    total_rain = round(sum(precipitation), 5),
    mean_ws = round(mean(wind_speed), 2),
    rain_duration = round(sum(rain_duration * 15 / 60), 2),
    mean_wd = round(circular.averaging(wind_direction), 2),
    mean_temp = round(mean(temperature), 2)
  )

# Filter rainless periods to calculate mean RH
weather_dat_no_rain <-
  read_excel(system.file(
    "extdata",
    "NC_weather_data.xlsx",
    package = "epiboxwoodblight",
    mustWork = TRUE
  )) %>%
  select(
    year,
    relative_humidity,
    leaf_wetness_duration,
    precipitation,
    location,
    spread_event,
    date
  ) %>%
  dplyr::mutate(year = as.factor(year)) %>%
  dplyr::mutate(location = as.factor(location)) %>%
  dplyr::mutate(spread_event = as.factor(spread_event)) %>%
  filter(precipitation == 0) %>%
  group_by(year, location, spread_event) %>%
  summarise(mean_rh = round(mean(relative_humidity * 100), 2))

# Combine data
weather_dat_comb <-
  left_join(weather_dat_rain,
            weather_dat_no_rain,
            by = c("year", "location", "spread_event"))

# Leaf wetness duration both inside and outside rainy periods
weather_wet <- read_excel(system.file(
  "extdata",
  "NC_weather_data.xlsx",
  package = "epiboxwoodblight",
  mustWork = TRUE
)) %>%
  dplyr::mutate(year = as.factor(year)) %>%
  dplyr::mutate(location = as.factor(location)) %>%
  dplyr::mutate(spread_event = as.factor(spread_event)) %>%
  group_by(year, location, spread_event) %>%
  summarise(lwd_duration = round(sum(leaf_wetness_duration / 60), 2))

weather_dat <-
  left_join(weather_dat_comb,
            weather_wet,
            by = c("year", "location", "spread_event"))

# Divide week 1 of 2014 rain/rain duration/wetness duration by 4 & that of week 2 & 3 by 3 to convert to per week data because the duration of spread event was 4 and 3 weeks, respectively.

weather_dat <- weather_dat %>%
  mutate(
    total_rain = ifelse(
      year == "2017" & spread_event == "1",
      total_rain / 4,
      ifelse(
        year == "2017" &
          spread_event %in% c("2", "3"),
        total_rain / 3,
        total_rain
      )
    ),
    rain_duration = ifelse(
      year == "2017" & spread_event == "1",
      rain_duration / 4,
      ifelse(
        year == "2017" &
          spread_event %in% c("2", "3"),
        rain_duration / 3,
        rain_duration
      )
    ),
    lwd_duration = ifelse(
      year == "2017" & spread_event == "1",
      lwd_duration / 4,
      ifelse(
        year == "2017" &
          spread_event %in% c("2", "3"),
        lwd_duration / 3,
        lwd_duration
      )
    )
  )
```



### Cobmine weather & disease data

Combine weather and disease data

```{r combine-disease-weather-data, warning=FALSE, message=FALSE}
dat_NC <-
  left_join(disease_dat,
            weather_dat,
            by = c("year", "location", "spread_event")) %>%
  # Replace NA with zero because NA are introduced due to data munging. Original values were zero
  dplyr::mutate(total_rain = replace_na(total_rain, 0)) %>%
  dplyr::mutate(rain_duration = replace_na(rain_duration, 0))

# Since we filtered data separately for precipitation and then without precipitation, NAs are introduced. In this step, data (in which values were added manually) is imported

dat_missing <- read_excel(system.file(
  "extdata",
  "NC_missing_data.xlsx",
  package = "epiboxwoodblight",
  mustWork = TRUE
)) %>%
  dplyr::mutate(year = as.factor(year)) %>%
  dplyr::mutate(location = as.factor(location)) %>%
  dplyr::mutate(spread_event = as.factor(spread_event))

# Combine data to replace NA values with distinct data
dat_nc <-
  left_join(dat_NC, dat_missing, by = c("year", "location", "spread_event")) %>%
  mutate(mean_ws = coalesce(mean_ws.x, mean_ws.y)) %>%
  select(-mean_ws.x, -mean_ws.y) %>%
  mutate(mean_temp = coalesce(mean_temp.x, mean_temp.y)) %>%
  select(-mean_temp.x, -mean_temp.y) %>%
  mutate(mean_rh = coalesce(mean_rh.x, mean_rh.y)) %>%
  select(-mean_rh.x, -mean_rh.y) %>%
  mutate(mean_wd = coalesce(mean_wd.x, mean_wd.y)) %>%
  select(-mean_wd.x, -mean_wd.y) %>%
  mutate(lwd_duration = coalesce(lwd_duration.x, lwd_duration.y)) %>%
  select(-lwd_duration.x, -lwd_duration.y) %>%
  distinct()

dat_nc <- dat_nc %>%
  mutate(daily_rain = round(total_rain/spev_duration, 2),
         daily_lwd  = round(lwd_duration/spev_duration, 2))



# Filter out mulch treatment. Use non-mulch and CP only.
dat_nc_ncb <- dat_nc %>%
  filter(treatment != "mulch", treatment != "between_row") # filter non-mulch, CP and between row treatments data

# Data considering only CP treatment
dat_cp <- dat_nc %>%
  filter(treatment == "CP")

# Data considering only leaf debris treatment
dat_ld <- dat_nc %>%
  filter(treatment == "non_mulch")

# Data considering only between row treatment
dat_br <- dat_nc %>%
  filter(treatment == "between_row")

# Data for Lambsburg site only
dat_lambsburg <- dat_nc %>%
  filter(location == "Lambsburg") %>%
  filter(treatment != "mulch", treatment != "between_row")
```



## Boxwood blight infection risk model

Run [Boxwood blight infection risk model](https://uspest.org/wea/Boxwood_blight_model_instructions.pdf). This the SENSOR-ADJ version of the model which uses leaf wetness recorded by the leaf wetness sensor
```{r run-model}
base_temperature <- 44


lookup_table <- read_excel(system.file(
  "extdata",
  "lookup_table.xlsx",
  package = "epiboxwoodblight",
  mustWork = TRUE
)) 
  

# WARNING: Double check for index_new (lookup_table) and index_anton (same but proportional wetness) 

index <- read_excel(system.file(
  "extdata",
  "hourly_df_index_new.xlsx",
  package = "epiboxwoodblight",
  mustWork = TRUE
)) %>%
 mutate(dry_hour_count = ifelse(lwd_duration == 0, sequence(rle(lwd_duration == 0)$lengths), 0)) %>%
  mutate(mean_temp_f = round(mean_temp_f)) %>%
  mutate(zero_index = lwd_duration == 0 | dry_hour_count > 5 | mean_temp_f < 44 | mean_temp_f > 86) %>%
  group_by(year, location, spread_event, cumsum(dry_hour_count > 5)) %>%
  # model version 1
  mutate(index_new = cumsum(ifelse(zero_index, 0, mean_temp_f - base_temperature))) %>% 
  relocate(index, .before = index_new) %>%
  left_join(lookup_table, by = "mean_temp_f") %>%
  # model version 3
  mutate(index_new = cumsum(ifelse(zero_index, 0, index_lookup))) %>% 
  mutate(index_new_anton = cumsum(ifelse(zero_index, 0, index_lookup) * lwd_duration / 60)) %>% # Proportional leaf wetness suggested by Anton
  mutate(dhs_exceeding_250_anton = ifelse(index_new_anton >= 250, TRUE, FALSE),
         index_rs_hour = ifelse(index_new_anton > 0, index_new_anton - lag(index_new_anton, default = 0), 0)) %>%
  select(-`cumsum(dry_hour_count > 5)`, -zero_index) 


# Find maximum value of index and whether disease is predicted or not, and combine with the data 
index_cap <- index %>%
  group_by(year, location, spread_event) %>%
  summarise(max_index_value_rs = round(max(index_new_anton)),
            disease_predicted_rs = ifelse(any(index_new_anton>= 250), 1, 0),
             index_rs_week = round(sum(index_rs_hour)))

  
index <- left_join(index, index_cap, by = c("year", "location", "spread_event")) %>%
    mutate(year = as.factor(year), spread_event = as.factor(spread_event)) 

# Finally, combine disease data with the index table 

# Disease data
ncncb <- dat_nc_ncb %>%
  select(spread_event, year, location, total_count) %>%
  mutate(year = as.factor(year), spread_event = as.factor(spread_event)) %>%
  group_by(year, location, spread_event) %>%
  summarise(total_count_index = sum(total_count),
            epidemic_week = ifelse(total_count>298, "Yes", "No"))



index <- left_join(index, ncncb, by = c("year", "location", "spread_event")) %>%
   mutate(year = as.factor(year), spread_event = as.factor(spread_event)) 
```



## Model comparison

Compare the model using three different leaf wetness criteria. 

`lws` represents `SENSOR` model
`rs` represents `SENSOR-ADJ` model
`nlws` represents `ESTIM-LW` model

```{r compare-models, message=FALSE, warning=FALSE}
# Import data and calculate hourly and weekly index
comparison_dat1 <- read_excel(
  system.file(
    "extdata",
    "Indices_validation_data_new.xlsx",
    package = "epiboxwoodblight",
    mustWork = TRUE
  )
) %>%
  mutate(year = as.factor(year)) %>%
  mutate(location = as.factor(location)) %>%
  mutate(week = as.factor(week)) %>%
  group_by(year, location, week) %>%
  mutate(
    index_lws_hour = ifelse(index_lws > 0, index_lws - lag(index_lws, default = 0), 0),
    index_lws_week = sum(index_lws_hour),
    disease_predicted_lws = ifelse(any(index_lws >= 250), "Yes", "No"),
    max_index_value_lws = max(index_lws),
    index_nlws_hour = ifelse(index_nlws > 0, index_nlws - lag(index_nlws, default = 0), 0),
    disease_predicted_nlws = ifelse(any(index_nlws >= 250), "Yes", "No"),
    max_index_value_nlws = max(index_nlws),
    index_nlws_week = sum(index_nlws_hour)
  ) %>%
  ungroup()

#openxlsx::write.xlsx(comparison_dat1, "indices_lc.xlsx")

comparison_dat2 <- comparison_dat1 %>%
  group_by(year, location, week) %>%
  summarise(
    disease_predicted_lws = ifelse(any(index_lws >= 250), 1, 0),
    max_index_value_lws = max(index_lws),
    max_index_value_nlws = max(index_nlws),
    index_lws_week = sum(index_lws_hour),
    index_nlws_week = sum(index_nlws_hour),
    disease_predicted_nlws = ifelse(any(index_nlws >= 250), 1, 0)
  ) %>%
  ungroup()


# Disease data
ncncb <- dat_nc %>%
  filter(treatment != "mulch") %>%
  rename(week = spread_event) %>%
  group_by(year, location, week) %>%
  summarise(
    disease_recorded_detector_plants = ifelse(any(total_count > 0), 1, 0),
    total_count = sum(total_count)
  )


comparison_dat3 <-
  left_join(comparison_dat2, ncncb, by = c("year", "location", "week"))

# Join index that was re-started each week
index_rs <- index_cap %>%
  rename(week = spread_event) %>%
  mutate(year = as.factor(year),
         week = as.factor(week))

comparison_dat4 <-
  left_join(comparison_dat3, index_rs, by = c("year", "location", "week"))


```



## Totat weekly accumulated blight risk index & infected leaf count relationship

### GLMs 

```{r leaf-count-model-association}
set.seed(42)

dat_2014 <- comparison_dat4 %>%
  filter(year==2014)

dat_2015 <- comparison_dat4 %>%
  filter(year==2015)

dat_2016 <- comparison_dat4 %>%
  filter(year==2016)

dat_2017 <- comparison_dat4 %>%
  filter(year==2017)

# Fit & plot LWS models for each year
lws_2014 <- glm(total_count ~ index_lws_week, family = quasipoisson, data= dat_2014)

summary(lws_2014)

#calculate McFadden's R-squared for model
r_squared_lws_2014 <- with(summary(lws_2014), 1 - deviance/null.deviance)


lws_2015 <- glm(total_count ~ index_lws_week, family = quasipoisson, data= dat_2015)

summary(lws_2015)

#calculate McFadden's R-squared for model
r_squared_lws_2015 <- with(summary(lws_2015), 1 - deviance/null.deviance)



lws_2016 <- glm(total_count ~ index_lws_week, family = quasipoisson, data= dat_2016)

summary(lws_2016)

#calculate McFadden's R-squared for model
r_squared_lws_2016 <- with(summary(lws_2016), 1 - deviance/null.deviance)


lws_2017 <- glm(total_count ~ index_lws_week, family = quasipoisson, data= dat_2017)

summary(lws_2017)

#calculate McFadden's R-squared for model
r_squared_lws_2017 <- with(summary(lws_2017), 1 - deviance/null.deviance)

# plot all LWS models

Lws_1 <-
  plot(ggpredict(lws_2014, "index_lws_week"), rawdata = TRUE) +
  labs(x = "", y = "Number of infected leaves", title = 2014) +
  coord_cartesian(ylim = c(0, 1000), xlim = c(0, 3000)) +
   annotate("text",
           x = c(750, 750),
           y = c(700, 800),
           label =c("R²= 0.37", "p=0.0114")) +
  theme_few() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank()) +
 theme(aspect.ratio = 1)



Lws_2 <-
  plot(ggpredict(lws_2015, "index_lws_week"), rawdata = TRUE) +
  labs(x = "", y = "", title = 2015) +
  coord_cartesian(ylim = c(0, 1000), xlim = c(0, 3000)) +
   annotate("text",
           x = c(750, 750),
           y = c(700, 800),
           label =c("R²= 0.47", "p=0.0006")) +
  theme_few() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.text.y = element_blank()) +
  theme(aspect.ratio = 1)



Lws_3 <-
  plot(ggpredict(lws_2016, "index_lws_week"), rawdata = TRUE) +
  labs(x = "LWS model", y = "", title = 2016) +
  coord_cartesian(ylim=c(0, 1000),xlim = c(0, 3000)) +
  annotate("text",
           x = c(750, 750),
           y = c(700, 800),
           label =c("R²= 0.02", "p=0.6283")) +
  theme_few() +
  theme(plot.title = element_text(hjust = 0.5)) +
 theme(aspect.ratio = 1)




Lws_4 <-
  plot(ggpredict(lws_2017, "index_lws_week"), rawdata = TRUE) +
  labs(x = "LWS model", y = "", title = 2017) +
  coord_cartesian(ylim=c(0, 1000),xlim = c(0, 3000)) +
  annotate("text",
           x = c(750, 750),
           y = c(700, 800),
           label =c("R²= 0.3", "p=0.0338")) +
  theme_few() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.y = element_blank()) +
   theme(aspect.ratio = 1)
  




lws_plots <- (Lws_1 +Lws_2)/ (Lws_3+ Lws_4) + 
  plot_layout(widths = c(1, 1), heights = c(1, 1))


# Fit and plot NLWS models for each year & then combine plots for all four years

nlws_2014 <- glm(total_count ~ index_nlws_week, family = quasipoisson, data= dat_2014)

summary(nlws_2014)

#calculate McFadden's R-squared for model
r_squared_nlws_2014 <- with(summary(nlws_2014), 1 - deviance/null.deviance)


nlws_2015 <- glm(total_count ~ index_nlws_week, family = quasipoisson, data= dat_2015)

summary(nlws_2015)

#calculate McFadden's R-squared for model
r_squared_nlws_2015 <- with(summary(nlws_2015), 1 - deviance/null.deviance)



nlws_2016 <- glm(total_count ~ index_nlws_week, family = quasipoisson, data= dat_2016)

summary(nlws_2016)

#calculate McFadden's R-squared for model
r_squared_nlws_2016 <- with(summary(nlws_2016), 1 - deviance/null.deviance)


nlws_2017 <- glm(total_count ~ index_nlws_week, family = quasipoisson, data= dat_2017)

summary(nlws_2017)

#calculate McFadden's R-squared for model
r_squared_nlws_2017 <- with(summary(nlws_2017), 1 - deviance/null.deviance)

# plot all NLWS models
Nlws_1 <-
  plot(ggpredict(nlws_2014, "index_nlws_week"), rawdata = TRUE) +
  labs(x = "", y = "", title = 2014) +
  coord_cartesian(ylim = c(0, 1000), xlim = c(0, 3000)) +
   annotate("text",
           x = c(750, 750),
           y = c(700, 800),
           label =c("R²= 0.13", "p=0.2393")) +
  theme_few() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.text.y = element_blank()) +
     theme(aspect.ratio = 1)



Nlws_2 <-
  plot(ggpredict(nlws_2015, "index_nlws_week"), rawdata = TRUE) +
  labs(x = "", y = "", title = 2015) +
  coord_cartesian(ylim = c(0, 1000), xlim = c(0, 3000)) +
   annotate("text",
           x = c(750, 750),
           y = c(700, 800),
           label =c("R²= 0.25", "p=0.0247")) +
  theme_few() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.text.y = element_blank()) +
     theme(aspect.ratio = 1)



Nlws_3 <-
  plot(ggpredict(nlws_2016, "index_nlws_week"), rawdata = TRUE) +
  labs(x = "NLWS model", y = "", title = 2016) +
  coord_cartesian(ylim=c(0, 1000),xlim = c(0, 3000)) +
  annotate("text",
           x = c(750, 750),
           y = c(700, 800),
           label =c("R²= 0.03", "p=0.5884")) +
  theme_few() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.y = element_blank()) +
     theme(aspect.ratio = 1)



Nlws_4 <-
  plot(ggpredict(nlws_2017, "index_nlws_week"), rawdata = TRUE) +
  labs(x = "NLWS model", y = "", title = 2017) +
  coord_cartesian(ylim=c(0, 1000),xlim = c(0, 3000)) +
  annotate("text",
           x = c(750, 750),
           y = c(700, 800),
           label =c("R²= 0.23", "p=0.0703")) +
  theme_few() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.y = element_blank()) +
     theme(aspect.ratio = 1)



Nlws_plots <- (Nlws_1 +Nlws_2)/ (Nlws_3+ Nlws_4) +
  plot_layout(widths = c(1, 1), heights = c(1, 1))

# Nlws_plots <- annotate_figure(Nlws_plots, left = textGrob("Number of infected leaves", rot = 90, vjust = 1), bottom = textGrob(""))




# Fit and plot RS models for each year & then combine plots for all four years

rs_2014 <- glm(total_count ~ index_rs_week, family = quasipoisson, data= dat_2014)

summary(rs_2014)

#calculate McFadden's R-squared for model
r_squared_rs_2014 <- with(summary(rs_2014), 1 - deviance/null.deviance)


rs_2015 <- glm(total_count ~ index_rs_week, family = quasipoisson, data= dat_2015)

summary(rs_2015)

#calculate McFadden's R-squared for model
r_squared_rs_2015 <- with(summary(rs_2015), 1 - deviance/null.deviance)



rs_2016 <- glm(total_count ~ index_rs_week, family = quasipoisson, data= dat_2016)

summary(rs_2016)

#calculate McFadden's R-squared for model
r_squared_rs_2016 <- with(summary(rs_2016), 1 - deviance/null.deviance)


rs_2017 <- glm(total_count ~ index_rs_week, family = quasipoisson, data= dat_2017)

summary(rs_2017)

#calculate McFadden's R-squared for model
r_squared_rs_2017 <- with(summary(rs_2017), 1 - deviance/null.deviance)

# plot all NLWS models
rs_1 <-
  plot(ggpredict(rs_2014, "index_rs_week"), rawdata = TRUE) +
  labs(x = "", y = "", title = 2014) +
  coord_cartesian(ylim = c(0, 1000), xlim = c(0, 3000)) +
   annotate("text",
           x = c(750, 750),
           y = c(700, 800),
           label =c("R²= 0.39", "p=0.0092")) +
  theme_few() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.text.y = element_blank()) +
      theme(aspect.ratio = 1)



rs_2 <-
  plot(ggpredict(rs_2015, "index_rs_week"), rawdata = TRUE) +
  labs(x = "", y = "", title = 2015) +
  coord_cartesian(ylim = c(0, 1000), xlim = c(0, 3000)) +
   annotate("text",
           x = c(750, 750),
           y = c(700, 800),
           label =c("R²= 0.48", "p=0.0003")) +
  theme_few() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.text.y = element_blank()) +
      theme(aspect.ratio = 1)



rs_3 <-
  plot(ggpredict(rs_2016, "index_rs_week"), rawdata = TRUE) +
  labs(x = "RS model", y = "", title = 2016) +
  coord_cartesian(ylim=c(0, 1000),xlim = c(0, 3000)) +
  annotate("text",
           x = c(750, 750),
           y = c(700, 800),
           label =c("R²= 0.02", "p=0.6028")) +
  theme_few() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.y = element_blank()) +
      theme(aspect.ratio = 1)




rs_4 <-
  plot(ggpredict(rs_2017, "index_rs_week"), rawdata = TRUE) +
  labs(x = "RS model", y = "", title = 2017) +
  coord_cartesian(ylim=c(0, 1000),xlim = c(0, 3000)) +
  annotate("text",
           x = c(750, 750),
           y = c(700, 800),
           label =c("R²= 0.38", "p=0.0132")) +
  theme_few() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.y = element_blank()) +
      theme(aspect.ratio = 1)



rs_plots <- (rs_1 + rs_2)/ (rs_3+ rs_4) +
  plot_layout(widths = c(1, 1), heights = c(1, 1))

# rs_plots <- annotate_figure(rs_plots, left = textGrob("Number of infected leaves", rot = 90, vjust = 1))


fig_index_lesions <- wrap_plots(lws_plots, Nlws_plots, rs_plots)

  
  ggsave(
    here("man", "figures/fig_index_lesions.png"),
    plot = fig_index_lesions,
    width = 14,
    height = 6,
    units = "in",
    dpi = 600
  )
  
fig_index_lesions  
```



### Plot total weekly accumulated blight risk index values 

```{r plot-hourly-infection-risk-index, warning=FALSE}
dat_weeeek <- dat_nc_ncb %>%
  mutate(date_in = format(date_in, "%m/%d")) %>%
  group_by(year, date_in) %>%
  arrange(year, date_in, total_count) %>%
  ungroup() %>%
  mutate(date_in = as.factor(date_in)) %>%
  mutate(treatment = recode(treatment,
                            "CP" = "Infected canopies",
                            "non_mulch" = "Leaf debris")) %>%
  rename("Inoculum source" = treatment) %>%
  select(date_in,
         year,
         location,
         spread_event,
         total_count,
         `Inoculum source`)




comparison_dat_6 <- comparison_dat4 %>%
  rename(spread_event = week) %>%
  pivot_longer(
    cols = c("index_lws_week",
             "index_nlws_week",
             "index_rs_week"),
    names_to = "Model",
    values_to = "index_weekly_value"
  )

index_dfff <- comparison_dat_6 %>%
  mutate(
    Model = recode(
      Model,
      "index_rs_week" = "SENSOR-ADJ model",
      "index_lws_week" = "SENSOR model",
      "index_nlws_week" = "ESTIM-LW model"
    )
  )


dat_week <-
  left_join(dat_weeeek,
            index_dfff,
            by = c("year", "location", "spread_event")) %>%
  distinct(year, location, spread_event, Model, .keep_all = TRUE)

openxlsx::write.xlsx(dat_week, "check.xlsx")

dates_weeks <- dat_week |>
  mutate(date = mdy(paste0(date_in, year)),
         week = week(date)) %>%
  mutate(week = ifelse(year == 2015 & spread_event == 24, 43, week))



min_week <- min(dates_weeks$week)
max_week <- max(dates_weeks$week)

year_breaks <- 2017:2014 |>
  set_names() |>
  map(\(year_n) {
    year_dat <- dates_weeks |>
      filter(year == year_n) |>
      select(date_in, week) |>
      unique()
    
    week_map <- character(max_week - min_week + 1)
    names(week_map) <- seq(min_week, max_week)
    
    week_map[as.character(year_dat$week)] <-
      as.character(year_dat$date_in)
    
    
    week_map
    
  })


p <- dates_weeks |>
  mutate(week = factor(week),
         `Inoculum source` = factor(`Inoculum source`)) |>
  nest(data = -year) |>
  mutate(pl = map2(data, year, \(data, year) {
    # browser()
    
    ggplot(data, aes(x = week, y = index_weekly_value, fill = Model)) +
      geom_col(position = position_dodge(width = 0.7)) +
      coord_cartesian(ylim = c(4, 8603)) +
      # scale_y_continuous(breaks = seq(0, 334305, length.out = 5), limits = c(0, 334305),
      #                      labels = function(x) sprintf("%.0f", x)) +
      scale_x_discrete(labels = year_breaks[[as.character(year)]],
                       drop = FALSE) +
      facet_wrap(year, scales = "free_y", strip.position = "right") +
      #  scale_fill_manual(values = c(
      #   "max_index_value_rs"= "navyblue",
      #   "max_index_value_lws" = "green",
      #    "max_index_value_nlws"= "darkred"
      # )) +
      scale_fill_viridis_d(option = "cividis") +
      theme_few(base_size = 10)
    
    
  })) |>
  pull(pl) |>
  wrap_plots(ncol = 1) +
  plot_layout(guides = "collect") & xlab(NULL) & ylab(NULL) &
  theme(
    legend.position = "top",
    axis.text.x = element_text(
      angle = 90,
      hjust = 1,
      vjust = 0.5,
      color = "black"
    ),
    axis.text.y = element_text(vjust = 0.5, color = "black"),
    axis.title = element_text(color = "black"),
    strip.text = element_text(face = "bold", color = "black")
    
  )


p <-
  p + labs(x = "Date of detector plants placement in the field") +
  theme(axis.title.x = element_text(hjust = 0.5))  # Center x-axis label


p_sum_index <- wrap_elements(p) +
  labs(tag = "Total weekly accumulated blight risk index") +
  theme(plot.tag = element_text(size = rel(1), angle = 90),
        plot.tag.position = "left")



ggsave(
  here("man", "figures/p_sum_index.png"),
  plot = p_sum_index,
  width = 7,
  height = 7,
  units = "in",
  dpi = 1000
)

ggsave(
  here("man", "figures/p_sum_index.eps"),
  plot = p_sum_index,
  width = 7,
  height = 7,
  units = "in",
  dpi = 600,
  device =  cairo_ps
)

p_sum_index
```



## Quantitative comparison of the model using different leaf wetness criteria

### Predictions when leaf wetness is recorded by leaf wetness sensor

### Confusion matrices 

```{r confusion-matrices}
disease_predicted_lws <-
  factor(comparison_dat4$disease_predicted_lws)
disease_predicted_nlws <-
  factor(comparison_dat4$disease_predicted_nlws)
disease_predicted_rs <- factor(comparison_dat4$disease_predicted_rs)
disease_recorded <-
  factor(comparison_dat4$disease_recorded_detector_plants)

confusion_matrix_lws <-
  metrica::confusion_matrix(
    obs = disease_recorded,
    pred = disease_predicted_lws,
    plot = TRUE,
    print_metrics = TRUE,
    metrics_list = c("accuracy", "precision", "recall", "specificity"),
    position_metrics = "top",
    colors = c(low = "#1f78b4", high = "#33a02c"),
    na.rm = TRUE
  )

confusion_matrix_nlws <-
  metrica::confusion_matrix(
    obs = disease_recorded,
    pred = disease_predicted_nlws,
    plot = TRUE,
    print_metrics = TRUE,
    metrics_list = c("accuracy", "precision", "recall", "specificity"),
    position_metrics = "top",
    colors = c(low = "#1f78b4", high = "#33a02c"),
    na.rm = TRUE
  )


confusion_matrix_rs <-
  metrica::confusion_matrix(
    obs = disease_recorded,
    pred = disease_predicted_rs,
    plot = TRUE,
    print_metrics = TRUE,
    metrics_list = c("accuracy", "precision", "recall", "specificity"),
    position_metrics = "top",
    colors = c(low = "#1f78b4", high = "#33a02c"),
    na.rm = TRUE
  )


fig_matrix <-
  confusion_matrix_lws + confusion_matrix_nlws + confusion_matrix_rs + plot_layout(ncol =
                                                                                     1, tag_level = 'new') +
  plot_annotation(tag_levels = list(c('(A)', '(B)', '(C)')))


fig_matrix
```


### F1 scores

```{r F1-Score}
# Calculate accuracy
accuracy_rs <- mean(disease_predicted_rs == disease_recorded)
# Calculate precision
precision_rs <-
  sum(disease_predicted_rs == 1 &
        disease_recorded == 1) / sum(disease_predicted_rs == 1)
# Calculate recall
recall_rs <-
  sum(disease_predicted_rs == 1 &
        disease_recorded == 1) / sum(disease_recorded == 1)

# Calculate F1-score
f1_score_rs <-
  round(2 * (precision_rs * recall_rs) / (precision_rs + recall_rs), 2)
```



### Reciever Operating Characteristics Curve (ROC)

```{r roc}
roc_lws <-
  pROC::roc(as.integer(disease_predicted_lws),
            as.integer(disease_recorded))
roc_plot_lws <-
  plot(roc_lws, main = "ROC curve for index_lws", col = "blue")
abline(0, 1, lty = 2, col = "gray")

roc_nlws <-
  pROC::roc(as.integer(disease_predicted_nlws),
            as.integer(disease_recorded))
roc_plot_nlws <-
  plot(roc_nlws, main = "ROC curve for index_nlws", col = "blue")
abline(0, 1, lty = 2, col = "gray")

roc_rs <-
  pROC::roc(as.integer(disease_predicted_rs),
            as.integer(disease_recorded))
roc_plot_rs <-
  plot(roc_rs, main = "ROC curve for index_rx", col = "blue")
abline(0, 1, lty = 2, col = "gray")

```







